To Airbnb?: A Question of Revenue

### 2018-07-23 review/response_to_reviewers.Rmd

---
title: "Responses to Reviewers for: To Airbnb?: A Question of Revenue"
#subtitle: "Andy Krause and Gideon Aschwanden"
#author: "Greenfield Advisors and University of Melbourne"
date: "July 22nd, 2018"
output:
  pdf_document: default
  html_document: default
header-includes: 
   - \usepackage{setspace}\doublespacing
   - \usepackage{caption}
   - \captionsetup[table]{skip=10pt}
geometry: margin=3.3cm
editor_options: 
  chunk_output_type: console
---

@@

* Regarding the structure of cities and the generalizability of the findings.  We agree that Melbourne, as a world class tourist destination and (arguably) the world's most livable city, does provide an example that does not necessarily translate well to a second or third tier city in many other parts of the world. As you noted, this desire for both tourists and residents to occupy the same space is what, in our minds, made this the ideal city for this analysis. We do feel that many of the world's first class cities are experiencing the same phenomenon -- namely that a large segment of the local population desires the same walkable, vibrant locations that tourists do, thus leading to increased demand and increasingly contentious discussions over the negative externalities of a thriving short-term rental market.  We've added some language in the beginning of the 'Limitations' section that addresses the issues of generalizability.  Finally, we are in the process of additional data collection to expand the research to a more representative sample.

## Reviewer #2

1. *The authors state “And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.” I agree with this statement. However, if the majority of the properties are vacation homes that owners rent out when they are not using them personally, then the hypothetical exercise performed in the paper is for naught. By definition, the owners cannot use a LTR strategy since it would preclude personal use throughout the year. In which case, the question “To Airbnb?” is not a question at all.*

* We agree that this exercise wouldn't apply to an owner who leases out their home while away but occupies it at other times.  If an investor were to buy that home they would have the choice of doing either an LTR or an STR which is why we framed the hypothetical situation as that of an investor. In the sensitivity analysis we attempted to address this potential bias in the sample by separately looking at those properties which were listed as Available or Reserved (not Blocked) at least 75% of the time (Tables 5 and 6).  We believe that these are clearly situations where it is not an owner-occupied who is being opportunistic, but rather a profit-seeking owner/investor.  For context, the "Opportunistic Sharers", those owners who only list their home very periodically (less than 25\% of the time), make up only 18% of all owners in our sample.

*This is an issue as the authors state “Beach properties are, relatively speaking, over-represented in the short-term market."* *Given the over-representation of beach properties, the differing STR and LTR revenues may be a product of the property-type (vacation vs. non-vacation) or a clientele effect (vacation homeowners vs. investor).* *The owner of a vacation home rents their property to subsidize its cost when they are not consuming its services personally. Thus, the only choice for the vacation home is the STR strategy. Short-term vacation rentals existed prior to Airbnb, so this is not a new phenomenon.*

* It is true that beach properties (mostly beach apartments) are over-represented in the STR data.  As are city-core apartments. In a situation with much greater quantities of data and the ability to match exact properties from STR and LTR markets, a matching approach would be the preferred method to eliminate these sample differences. We also considered some down-sampling of these two over-represented classes, but felt that further restricting an already small sample was not advantageous.  What we have done to limit the impact of these differences is three fold:

A) We have limited the study area (along the beach) to the 'urban' area of Melbourne where employment opportunities abound and where long-term renters are likely to demand to live.  More simply, we have not considered homes in the Mornington Peninsula, the traditional summer beach vacation area for Melburnians -- an area where there are many beach (and inland) homes whose highest and best use is heavily favored towards short term rentals (they are set up for large groups, include hot tubs, game rooms, extensive parking, etc. )

B) When estimating the long-term rental values we segment the models by submarket, so that we are only using beach properties to create beach rental values.  In this way, our comparison become a comparison of relative willingness to pay between short and long term renters for beach properties, not all properties.

C) We control for submarket type in the statistical models that we estimate.

Certainly, not all of the clientele effects are covered by these controls, but we feel that we have done what is available given the data quantity and quality.  The results, which show that beach properties have a considerably higher long-term preference rate, suggest that many of the obvious 'beach house'-type dwellings used by owners for vacations have been controlled for (or not included) in the study.

@@

3. *The short-term costs in Eq. 6 are understated. The authors note that short-term rentals are furnished by the owner, but do not include them as a cost. Given the feedback system built into Airbnb the furnishings may be a non-trivial expense. In other words, if the furnishing are in disrepair it may hurt the owners’ ratings and increase the vacancy rate of the STR. For this reason, I also expect the maintenance costs to be higher for the STR (especially since LTR has a security deposit to cover repairs that are not “normal wear and tear”). In addition, the authors do not address the question of who is managing the STR and responding to the Airbnb requests. Given the weekly (perhaps even nightly) turnover associated with STRs this is another cost that should be included/considered.*

 * As furnishings are a durable good, only some small percentage of the costs is truly experienced during the first year, the period this exercise examines.  While furniture is a bit like automobilies, the minute you sleep on a mattress, it loses a good bit of value, one year old furniture -- especially non-beds/sofa items do hold value relatively well (common rates of depreciation for tax purposes are 10% annually).  Thus, furnishing costs are borne over a long time period.  Further, we were not able to determine which of the LTR rentas are furnished so likely there is a bit of upward bias in those estimated values as well.  Neither of these facts eliminates the need to buy furniture to outfit an Airbnb, we simply did not have enough information on what the likely one year cost of furniture would be.  This is a relevant comment and we've tried to address its potential impact in the sensivity analysis discussed a few paragraphs below.

@@

 * We do agree that on-going (month by month) maintenance costs will likely be higher in an STR, however, large improvements, new carpets, painting, etc. do burden LTRs at tenant turnover.  As we mentioned on page 6, we believe there is an argument to be made that LTR tenants actually burden a property as much if not more by virtue of moving furniture in and out, hanging decorations and, likely, spending more actual time in the dwelling than a tourist would.  As a result of this and the fact that we were not able to find any evidence of greater maintenance costs on either side, we decided on assumming equivalent costs.

  * Certainly there are time costs of managing an STR.  These costs vary widely based on the owner's proximity to the STR and their individual imputed wage (the value of their time).  In general, these costs are somewhat idiosyncratic to the owner and we struggled with producing a standardized estimate.  On the converse, STR does allow for an early exit strategy, occupation or transfer to LTR on short notice, options which also have an unknown (or at least very owner-specific) value.  We opted to allow these two forces to 'balance' out and have not accounted for either.  The intent of the hypothetical exercise was to present the case of a simple investor (say a family) making one investment, not a large STR aggregation firm.  Note that there are firms in Melbourne and other cities that will manage your Airbnb for you.  Anecdotally, we've seen fees ranging from 8% to 15% for this service.

  * All of the above concerns could point to a potential underestimate of STR costs, which would lower STR net revenues.  As a further senitivity analysis, we reduced net STR revenues by 15% (10% for management fee and 5% for a furnishing costs), the results of which are found in the Table below. These should be compared to Table 2 in the paper.  Reducing STR revenues by 15% has the likely impact of decreasing the base rate of STR preference from 28% down to 22% (also evidenced in the intercept terms in the below table.)  The coefficients on the set of variables tested (while moving a bit) have generally remained stable in both terms of magnitude and direction.  We feel that your comments regarding management and furniture are worthwhile.  We worry about adding another table of sensitivity results to the paper in terms of confusion.  We have added additional language to the Discussion section to note these omissions, explain our rationale for doing so and discuss the sensitivity test shown in Table 1 of this response. We open to further addressing this issue if you feel it would be of assistance to readers.

@@

  # Print
  print(x.full,
       include.rownames=FALSE,
       hline.after=c(-1,0,16,21,21),
       # add.to.row = comment,
       caption.placement='top')

```

4. *The STR vs. LTR decision may not apply to every property. In the United States, some HOAs ban short-term rentals and impose substantial fines. Are there similar situations in Melbourne? Similarly, some cities in the United States have proposed legislature to ban STR if it is not the homeowner’s primary residence (example here). Are there similar situations in Australia?*

 * Very true.  The legality of short term rentals is very much in flux at the moment and varies from municipality to municipality, even building to building. One could imagine that in an area with high demand and a large number of dwellings being ineligible for STR leasing, that a premium might accrue to those that were eligible. At the time of this research, there were no governmental restrictions on short term use in Melbourne, though they were being hotly debated (which was the original impetus for this research).  Fines are now able to be levied for noise violations. currently, the State of Victoria has regulations that directly impact properties that allow more than five occupants (business license required). The City of Sydney has moved from being opposed to any regulations to now proposing that individual condominium buildings enact rules that they see fit.  Of course, rules can only be enforced if adequate data on occupancies is available.

5. *The abstract states “The growth of Airbnb and other short-term rental platforms have presented owners of residential properties with a choice of leasing strategy: traditional long-term rental or a short-term approach ("Airbnb-ing").” This option was available prior to Airbnb and the internet. Airbnb has “disrupted” the STR market, but it did not create it.*

@@

* Great point.  The short-term market for large, natural amenitiy-rich (ski + beach) properties is decades (if not centuries old).  We believe that Airbnb and its sister-sites have opened up this market to small, urban properties as well decreased search costs to the point that, from an owner perspective,  very short stays, even of a single night, can be economically feasible.  We have updated the abstract to represent the distinction that you have made.

*The next sentence in the abstract states “In this paper we identify those sitations -- location, structure type and property characteristics -- that lead to the highest likelihood of favoring a short-term strategy over a long term one.” I think sitations is a typo. I only point it out because it is in the abstract.*

### 2018-07-23 papers/full_paper.Rmd
Review: Finalizing responses to reviewers

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')
  setwd('..')
  #source(file.path(getwd(), 'functions', 'abb_Functions.R'))
  source(file.path('c:/code', 'research', 'airbnbMelbourne', 'functions', 'abb_Functions.R'))

@@

  # setwd('..')
  # load(file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))
  load(file.path('c:', 'code', 'research', 'airbnbMelbourne','data', 'analyzed',
                 'analyzed_data.RData'))

@@

The logistic regression models (Table 2, 4 and 6) than we estimate show that most of the finding from the univariate analyses done earlier are validated.  Small apartments in the central city are the best bet for generating greater revenues through short-term leasing.  We also test for the impact of host behavior on the probability of short-term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short-term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

@@

There are a number of limitations to this research.  First and foremost may be the generalizability of this work.  Melbourne is one of the primary examples of a world-class city in which local residents and tourists often (and fiercely) compete for the same urban spaces.  As such, the results we find here are likely to be much more applicable and generalizable to large, thriving cities with attractive urban neighborhoods such as London, San Francisco and Hong Kong, rather than secondary or tertiary metropolitan areas.  From investment and policy perspectives, however, it is in these larger centers that the findings of this research are most useful in a practical sense.

Next, there are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for long-term properties is that the short-term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short-term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short-term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.com.

@@

In this work, we have made the assumption that, over the course of a year, the general maintenance costs (excluding basic janitorial) are identical between short and long-term leasing options.  We have done so as we were unable to location any evidence to suggest that one was higher than the other. Additionally, in the reported results we have not made allowances for furniture and the turnover costs associated with managing an STR.  This omission is intentional as we felt these costs were highly idiosyncratic (turnover) and/or not limited to the one-year scenario period (furniture).  Sensitivity tests on these two costs (both of which lower STR revenues) do create decrease in the base rate of STR preference, but show little change to the coefficients across the various regression models. There is limited research around STR rental hosting costs and this uncertainty around cost estimates is a limitation of our study and could prove to be an avenue of future research.

Beyond Airbnb.com other short-term platforms exist (e.g. www.homeaway.com). This project doesn't take into account that individual properties may be advertised and rented out through different platforms. These are additional revenue streams that are not take into account and might increase the occupancy and revenue for short-term rentals. We do suggest that, due to block patterns, a number of host do appear to be employing this approach but cannot confirm the extend to which this is actually happening.

Finally, our analysis looks at a single year, beginning in September (spring) and ending in August in which the property owner is given a binary option of either short- or long-term leasing strategy.  In reality, the possibilities may not be so rigid.  For example,given the fact that Melbourne's large university student population (many of whom are international) require housing during Autumn, Winter and Spring, while tourists swell in the summer months, an enterprising owner maybe able to pair long-term leasing during the off-season with short-term leasing during peak times to maximize revenues.  In general, the rise of Airbnb and related portals give properties owners more options, and therefore more opportunities, to maximize returns on their property.  Future research should expand our scenario to look at more than just a single binary option over a fixed period.  We feel there may be significant opportunities for new ventures in industry to exploit the changing market for housing.

>>>>

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')
  #setwd('..')
  #source(file.path(getwd(), 'functions', 'abb_Functions.R'))
  source(file.path('c:/code', 'research', 'airbnbMelbourne', 'functions', 'abb_Functions.R'))

@@

  # setwd('..')
  # load(file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))
  load(file.path('c:/code', 'research', 'airbnbMelbourne','data', 'analyzed',
                 'analyzed_data.RData'))

@@

The logistic regression models (Table 2, 4 and 6) than we estimate show that most of the findings from the univariate analyses done earlier are validated.  Small apartments in the central city are the best bet for generating greater revenues through short-term leasing.  We also test for the impact of host behavior on the probability of short-term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short-term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

@@

There are a number of limitations to this research.  First, and foremost, may be the generalizability of this work.  Melbourne is one of the primary examples of a world-class city in which local residents and tourists often (and fiercely) compete for the same urban spaces.  As such, the results we find here are likely to be much more applicable and generalizable to large, thriving cities with attractive urban neighborhoods such as London, San Francisco and Hong Kong, rather than secondary or tertiary metropolitan areas.  From investment and policy perspectives, however, it is in these larger centers that the findings of this research are most useful in a practical sense.

Next, there are few shared property characterstics between the two datasets, which make cross-estimating rents and rates difficult, especially from the short-term market.  Part of the difficulty in estimating potential short-term rates for existing long-term properties is that the short-term market appears to be heavily influenced by non-property specific factors such as the leasing policies (minimum stay, number of guests, etc.) themselves.  Future research should focus both on building better models to explain (and ultimately predict) short-term rates and occupancies as well as to understand the impact of policies such as strict cancellation or no pets on the short-term market.  It is likely that this work is being done proprietarily by Airbnb.com itself and other complementary firms such as AirDNA.co.

@@

In this work, we have made the assumption that, over the course of a year, the general maintenance costs (excluding basic janitorial) are identical between short and long-term leasing options.  We have done so as we were unable to location any evidence to suggest that one was higher than the other. Additionally, in the reported results we have not made allowances for furniture and the turnover costs associated with managing an STR.  This omission is intentional as we felt these costs were highly idiosyncratic (turnover) and/or not limited to the one-year scenario period (furniture).  Sensitivity tests on these two costs (both of which lower STR revenues) do create decreases in the base rate of STR preference, but show little change to the coefficients across the various regression models. overall, there is limited research around STR rental hosting costs and this uncertainty around cost estimates is a limitation of our study and could prove to be an avenue of future research.

Beyond Airbnb.com other short-term platforms exist (e.g. www.homeaway.com). This project doesn't take into account that individual properties may be advertised and rented out through different platforms. These are additional revenue streams that are not take into account and might increase the occupancy and revenue for short-term rentals. We do suggest that, due to block patterns, a number of host do appear to be employing this approach but cannot confirm the extent to which this is actually happening.

Finally, our analysis looks at a single year, beginning in September (spring) and ending in August in which the property owner is given a binary option of either short- or long-term leasing strategy.  In reality, the possibilities may not be so rigid.  For example, given the fact that Melbourne's large university student population (many of whom are international) require housing during Autumn, Winter and Spring, while tourists swell in the summer months, an enterprising owner maybe able to pair long-term leasing during the off-season with short-term leasing during peak times to maximize revenues.  In general, the rise of Airbnb and related portals give properties owners more options, and therefore more opportunities, to maximize returns on their property.  Future research should expand our scenario to look at more than just a single binary option over a fixed period.  We feel there may be significant opportunities for new ventures in industry to exploit the changing market for housing.
